---
title: "week9assignment_RC"
output:
  pdf_document: default
  html_document: default
date: "2025-03-18"
---

## Set Up

```{r}

lizards <- read.csv("jrn_lizard.csv")

library(here)
library(tidyverse)
library(naniar)
library(DHARMa)
library(gtsummary)
library(broom)

lizards_filtered <- lizards %>%
  filter(spp == "UTST")

```

## Exploratory Figures

```{r}

hist1 <- ggplot(lizards_filtered, aes(x = SV_length)) +
  geom_histogram() +
  theme_bw()
hist1

bar1 <- ggplot(lizards_filtered, aes(x = tail)) + #cannot do histogram without continuous x
  geom_bar() +
  theme_bw()
bar1

bar2 <- ggplot(lizards_filtered, aes(x = sex)) + geom_bar() + theme_bw()
bar2

bar3 <- ggplot(lizards_filtered, aes(x = zone)) + geom_bar() + theme_bw()
bar3

```

## As Factors and Model

```{r}

unique(lizards_filtered$zone)

lizards_filtered$zone <- factor(lizards_filtered$zone,
                              levels = c("C", "G", "M", "T"))

unique(lizards_filtered$sex)

lizards_filtered$sex <- factor(lizards_filtered$sex,
                              levels = c("J", "F", "M"))

unique(lizards_filtered$tail)

lizards_filtered$tail <- factor(lizards_filtered$tail,
                              levels = c("B", "W"))

liz_mod <- glm(tail~SV_length+sex+zone,
                 data = lizards_filtered,
                 family = "binomial")

summary(liz_mod)

simulateResiduals(liz_mod) %>% plot()

```

The results of a logistic regression suggest that the log-odds of having a whole tail are greater with lower snout-to-vent length (B = -0.045, p < 0.001). There was also a significant decrease in log-odds of having a whole tail in vegetation zones with black grama grassland (B = -0.565, p = 0.001) and tarbush shrubland (B = -0.404, p = 0.025) relative to creosotebush shrubland. Mesquite duneland vegetation zones (B = -0.301, p = 0.063) did not have significant log-odds of having a whole tail relative to creosotebush shrubland. Neither females (B = -0.204, p = 0.322) nor males (B = -0.139, p = 0.534) had significant log-odds of having a whole tail relative to juveniles. 

## Predictive Figures

```{r, fig.cap="Figure 1. Predictive model of the probability of a lizard tail being whole in relation to snout-to-vent length by vegetation zone."}

# variables are SV length, sex, and zone

range(lizards_filtered$SV_length, na.rm=TRUE)

length_vector <- rep(seq(from = 5, to = 87), 12) #83, 3, and 4 don't have common factors

sex_vector <- rep(c("J", "F", "M"), 332)

zone_vector <- c(rep("C", 249),
                   rep("G", 249),
                 rep("M", 249),
                 rep("T", 249))

data_pred <- data.frame(sex_vector, length_vector, zone_vector)
colnames(data_pred) <- c("sex", "SV_length", "zone")

prediction <- predict(liz_mod,
                      newdata = data_pred,
                      type = "response",
                      se.fit = TRUE)


data_fig <- data.frame(data_pred,
                       prediction$fit,
                       prediction$se.fit)


colnames(data_fig) <- c("Sex", "Length", "Zone", "probability", "se")

ggplot(data_fig, aes(x = Length,
                     y = probability)) +
  geom_smooth(aes(color = Zone)) +
  geom_ribbon(aes(ymin = probability - se,
                  ymax = probability + se,
                  fill = Zone), alpha = 0.3) +
  labs(x = "Length", y = "Probability of Whole Tail", color = "Zone", fill = "Zone") +
  theme_bw()

# jagged because there are so many replicates?

```

```{r, fig.cap="Figure 2. Predictive model of the probability of a lizard tail being whole in relation to snout-to-vent length by sex."}

ggplot(data_fig, aes(x = Length,
                     y = probability)) +
  geom_smooth(aes(color = Sex)) +
  geom_ribbon(aes(ymin = probability - se,
                  ymax = probability + se,
                  fill = Sex), alpha = 0.3) +
  labs(x = "Length", y = "Probability of Whole Tail", color = "Sex", fill = "Sex") +
  theme_bw()

```

The probability of having a whole tail decreases as snout-to-vent length increases for all sexes and vegetation zones.

## Poisson Regression: Exploratory Plots

```{r}

lizards2 <- read_csv("jrn_lizard_npp.csv")

ggplot(lizards2, aes(x = BOER,
                     y = lizard_count)) +
  geom_point(aes(color = season)) +
  labs(x = "percent cover", y = "lizard count") +
  theme_bw()

ggplot(lizards2, aes(x = LATR,
                     y = lizard_count)) +
  geom_point(aes(color = season)) +
  labs(x = "percent cover", y = "lizard count") +
  theme_bw()

ggplot(lizards2, aes(x = PRGL,
                     y = lizard_count)) +
  geom_point(aes(color = season)) +
  labs(x = "percent cover", y = "lizard count") +
  theme_bw()

ggplot(lizards2, aes(x = SCBR,
                     y = lizard_count)) +
  geom_point(aes(color = season)) +
  labs(x = "percent cover", y = "lizard count") +
  theme_bw()

```

## Poisson Regression

```{r}

unique(lizards2$season)

lizards2$season <- factor(lizards2$season,
                             levels = c("F",
                                        "S",
                                        "W"))

lizard2_mod <- glm(lizard_count~season+BOER+LATR+PRGL+SCBR,
                data = lizards2,
                family = "poisson")

```

```{r}

summary(lizard2_mod)

plot(lizard2_mod)

lizards2_2 <- lizards2[-c(1,4,50),]

lizard2_mod_2 <- glm(lizard_count~season+BOER+LATR+PRGL+SCBR,
                 data = lizards2_2,
                 family = "poisson")

summary(lizard2_mod_2)

plot(lizard2_mod_2)

```

```{r}

lizards2_3 <- lizards2_2[-c(18),]

lizard2_mod_3 <- glm(lizard_count~season+BOER+LATR+PRGL+SCBR,
                 data=lizards2_3,
                 family = "poisson")

summary(lizard2_mod_3)

plot(lizard2_mod_3)

tbl_regression(lizard2_mod_3)

```

I removed a total of four outliers that were beyond Cook's distance.

The results of a logistic regression suggest that the log-odds of having higher lizard counts increase with burrograss vegetation (B = 0.490, p < 0.001). The log-odds of higher lizard counts decrease with increased black grama grass vegetation (B = -0.125, p < 0.001). There was a significant decrease in the log-odds of having higher lizard counts in the summer (B = -0.911, p < 0.001) relative to the fall. Neither increased creosote bush (B = 0.062, p = 0.575) nor honey mesquite (B = 0.090, p = 0.116) had significant log-odds of having higher lizard counts. Winter (B = 0.017, p = 0.847) did not have significant log-odds of having higher lizard counts relative to fall.

